Analytic reproducibility in articles receiving open data badges at Psychological Science: An observational assessment

### 2020-06-17 manuscript/manuscript.Rmd
minor text edits after MN feedback

---
title             : "Analytic reproducibility in articles receiving open data badges at Psychological Science: An observational assessment"
shorttitle        : "Analytic reproducibility"

@@

abstract: | 
  For any scientific report, repeating the original analyses upon the original data should yield the original outcomes. We evaluated analytic reproducibility in `r knitr::load_cache(label = 'abstract_text', object = 'totalArticles')` Psychological Science articles awarded open data badges. Initially, `r knitr::load_cache(label = 'abstract_text', object = 'intialReproRateArticles')` articles contained at least one “major numerical discrepancy” (>10% difference) prompting us to request input from original authors. Ultimately, target values were reproducible without author involvement for `r knitr::load_cache(label = 'abstract_text', object = 'repro_noAI')` articles; reproducible with author involvement for `r knitr::load_cache(label = 'abstract_text', object = 'repro_withAI')` articles; not fully reproducible with no substantive author response for `r knitr::load_cache(label = 'abstract_text', object = 'noRepro_noAI')` articles; and not fully reproducible despite author involvement for `r knitr::load_cache(label = 'abstract_text', object = 'noRepro_withAI')` articles. Overall, `r knitr::load_cache(label = 'abstract_text', object = 'majorErrors')` major numerical discrepancies remained out of `r knitr::load_cache(label = 'abstract_text', object = 'totalValues')` checked values `r knitr::load_cache(label = 'abstract_text', object = 'reproRateValues')`, but original conclusions did not appear affected. Non-reproducibility was primarily caused by unclear analysis reporting. Open data alone is not sufficient to ensure analytic reproducibility.

@@

Difficulty establishing the analytic reproducibility of research reports has been encountered in several scientific domains, including economics, political science, and psychology [@chang2015; @eubank2016; @stodden2018; @obels2019; @hardwicke2018a]. A preliminary obstacle for many such studies is that research data are typically unavailable [@hardwicke2018; @hardwicke2020a; @wicherts2006]. Even when data can be accessed, suboptimal management and inadequate documentation can complicate data reuse [@hardwicke2018a; @kidwell2016]. These failures to adequately preserve and share earlier stages of the research pipeline typically preclude downstream assessment of analytic reproducibility.

@@

In the present study, we intended to investigate whether the findings of @hardwicke2018a extended to a corpus of Psychological Science articles that received an ‘Open Data Badge’ to signal data availability. In order to focus specifically on the reproducibility of the analysis process rather than upstream data availability or management practices, we selected only articles that had reusable data according to a previous study [@kidwell2016]. The submission guidelines of Psychological Science specifically stated that authors could earn an Open Data badge for “making publicly available the digitally-shareable data necessary to reproduce the reported result.” [^psychsci1] Additionally, authors were asked to self-certify that they had provided “…sufficient information for an independent researcher to reproduce the reported results.” [^psychsci2] Thus, if this policy was operating as intended, all numbers presented in these articles should be independently reproducible. If not, we hoped to learn about causes of non-reproducibility in order to identify areas for improvement in journal policy and research practice.

@@

Of `r data_kidwell_badge %>% nrow()` articles marked with an open data badge, Kidwell and colleagues had identified `r data_kidwell_badge_reusable %>% nrow()` with datasets that met four reusability criteria (accessible, correct, complete, and understandable). For each of these articles, one investigator (TEH) attempted to identify a coherent set of descriptive and inferential statistics, roughly 2-3 paragraphs of text, sometimes including a table or figure, related to a ‘substantive’ finding based on ‘relatively straightforward’ analyses. In total, `r nrow(data_values)` discrete numerical values reported in `r totalArticles` articles were identified as target values. The articles were published between January 2014 and May 2015. Further information about the sample is available in Supplementary Information \@ref(sup_sample).

@@

```{r}
# prepare information on insufficent information errors for reporting inline
insuffInfoErrors <- data_articles %>% filter(Insufficient_Information_Errors > 0) %>% nrow() # how many articles had at least 1 insufficient information error
insuffInfoErrors_percent <- insuffInfoErrors/totalArticles*100 # calculate percent
insuffInfoErrors_CIs <- propCI(insuffInfoErrors,totalArticles) # caluclate CIs
```
Prior to author involvement, our reproducibility checks indicated that `r reproBefore %>% filter(reproOutcome == F) %>% pull(n)` out of the `r totalArticles` (`r reproBefore %>% filter(reproOutcome == F) %>% pull(percent)`) articles contained at least one major numerical discrepancy. After requesting input from original authors, several issues were resolved through the provision of additional information[^authorInfo] leaving `r reproAfter %>% filter(reproOutcome == F) %>% pull(n)` out of the `r totalArticles` (`r reproAfter %>% filter(reproOutcome == F) %>% pull(percent)`) articles containing at least one major numerical discrepancy. Even after author contact, we could not complete `r insuffInfoErrors` (`r insuffInfoErrors_percent`%, `r insuffInfoErrors_CIs`) reproducibility checks because insufficient information was available about aspects of the original analyses.

Ultimately, target values were fully reproducible for `r articleLevelSummary %>% filter(finalOutcome == 'Reproducible\nwithout author\ninvolvement') %>% pull(fullReport)` articles without requiring author involvement; fully reproducible for `r articleLevelSummary %>% filter(finalOutcome == 'Reproducible\nwith author\ninvolvement') %>% pull(fullReport)` articles only with author involvement; not fully reproducible for `r articleLevelSummary %>% filter(finalOutcome == 'Not reproducible\nwith author\ninvolvement') %>% pull(fullReport)` articles despite author involvement; and not fully reproducible for `r articleLevelSummary %>% filter(finalOutcome == 'Not reproducible\nwithout author\ninvolvement') %>% pull(fullReport)` articles with no substantive response from authors. Availability of original analysis scripts appeared to provide some modest benefits in terms of reproducibility outcomes and time expenditure (see Supplementary Information \@ref(sup_results)).

In no cases did the observed numerical discrepancies appear to be consequential for the conclusions stated in the original articles (see Supplementary Information \@ref(vignettes)).

@@

@@

Whilst the current findings are concerning, non-reproducibility is fortunately a solvable problem, at least in principle. As noted above, a major barrier to analytic reproducibility is low availability of research resources like data and code [@hardwicke2020a]. Some evidence suggests that journal policies which mandate data sharing can be highly effective at increasing data availability [@hardwicke2018a; @nuijten2017], apparently more so than voluntary badge schemes [@kidwell2016]. Sharing of analysis scripts may help by providing veridical documentation of the analysis process which is not well captured in verbal prose [@hardwicke2018a]. However, as with data sharing, mere availability will not be sufficient to confer the potential benefits of analysis scripts. The utility of scripts depends on a range of factors, including clarity, structure, and documentation, as well the programming language that is used and whether it is in a proprietary format or requires special expertise to understand. In the present study, script availability appeared to offer modest benefits in terms of reproducibility outcomes and time expenditure; however, as only six articles shared scripts[^psychsci_scripts], it is difficult to draw strong conclusions about their impact from this evidence alone.

@@

TEH and MCF designed the study. TEH, MB, KM, EH, MBN, BNP, BdM, BL, EJY, and MCF conducted reproducibility checks. TEH performed the data analysis. TEH and MCF wrote the manuscript. X provided feedback on the manuscript. All authors gave final approval for publication.

>>>>

---
title             : "Analytic reproducibility in articles receiving open data badges at Psychological Science: An observational study"
shorttitle        : "Analytic reproducibility"

@@

abstract: | 
  For any scientific report, repeating the original analyses upon the original data should yield the original outcomes. We evaluated analytic reproducibility in `r knitr::load_cache(label = 'abstract_text', object = 'totalArticles')` Psychological Science articles awarded open data badges between 2014-2015. Initially, `r knitr::load_cache(label = 'abstract_text', object = 'intialReproRateArticles')` articles contained at least one “major numerical discrepancy” (>10% difference) prompting us to request input from original authors. Ultimately, target values were reproducible without author involvement for `r knitr::load_cache(label = 'abstract_text', object = 'repro_noAI')` articles; reproducible with author involvement for `r knitr::load_cache(label = 'abstract_text', object = 'repro_withAI')` articles; not fully reproducible with no substantive author response for `r knitr::load_cache(label = 'abstract_text', object = 'noRepro_noAI')` articles; and not fully reproducible despite author involvement for `r knitr::load_cache(label = 'abstract_text', object = 'noRepro_withAI')` articles. Overall, `r knitr::load_cache(label = 'abstract_text', object = 'majorErrors')` major numerical discrepancies remained out of `r knitr::load_cache(label = 'abstract_text', object = 'totalValues')` checked values `r knitr::load_cache(label = 'abstract_text', object = 'reproRateValues')`, but original conclusions did not appear affected. Non-reproducibility was primarily caused by unclear analysis reporting. Open data alone is not sufficient to ensure analytic reproducibility.

@@

Difficulty establishing the analytic reproducibility of research reports has been encountered in several scientific domains, including economics, political science, and psychology [@chang2015; @eubank2016; @stodden2018; @obels2019; @hardwicke2018a; but see @naudet2018]. A preliminary obstacle for many such studies is that research data are typically unavailable [@hardwicke2018; @hardwicke2020a; @wicherts2006]. Even when data can be accessed, suboptimal management and inadequate documentation can complicate data reuse [@hardwicke2018a; @kidwell2016]. These failures to adequately preserve and share earlier stages of the research pipeline typically preclude downstream assessment of analytic reproducibility.

@@

In the present study, we intended to investigate whether the findings of @hardwicke2018a extended to a corpus of Psychological Science articles that received an ‘open data badge’ to signal data availability. In order to focus specifically on the reproducibility of the analysis process rather than upstream data availability or management practices, we selected only articles that had reusable data according to a previous study [@kidwell2016]. The submission guidelines of Psychological Science specifically stated that authors could earn an open data badge for “making publicly available the digitally-shareable data necessary to reproduce the reported result.” [^psychsci1] Additionally, authors were asked to self-certify that they had provided “…sufficient information for an independent researcher to reproduce the reported results.” [^psychsci2] Thus, if this policy was operating as intended, all numbers presented in these articles should be independently reproducible. If not, we hoped to learn about causes of non-reproducibility in order to identify areas for improvement in journal policy and research practice.

@@

Of `r data_kidwell_badge %>% nrow()` articles marked with an open data badge, Kidwell and colleagues had identified `r data_kidwell_badge_reusable %>% nrow()` with datasets that met four reusability criteria (accessible, correct, complete, and understandable). For each of these articles, one investigator (TEH) attempted to identify a coherent set of descriptive and inferential statistics, roughly 2-3 paragraphs of text, sometimes including a table or figure, related to a ‘substantive’ finding based on ‘relatively straightforward’ analyses. We focused on substantive findings because they are the most important and straightforward analyses to ensure that our team had sufficient expertise to re-run them. In total, `r nrow(data_values)` discrete numerical values reported in `r totalArticles` articles were identified as target values. The articles were published between January 2014 and May 2015. Further information about the sample is available in Supplementary Information \@ref(sup_sample).

@@

```{r}
# prepare information on insufficent information errors for reporting inline
insuffInfoErrors <- data_articles %>% filter(Insufficient_Information_Errors > 0) %>% nrow() # how many articles had at least 1 insufficient information error
insuffInfoErrors_percent <- insuffInfoErrors/totalArticles*100 # calculate percent
insuffInfoErrors_CIs <- propCI(insuffInfoErrors,totalArticles) # caluclate CIs
```
Prior to author involvement, our reproducibility checks indicated that `r reproBefore %>% filter(reproOutcome == F) %>% pull(n)` out of the `r totalArticles` (`r reproBefore %>% filter(reproOutcome == F) %>% pull(percent)`) articles contained at least one major numerical discrepancy. After requesting input from original authors, several issues were resolved through the provision of additional information[^authorInfo] leaving `r reproAfter %>% filter(reproOutcome == F) %>% pull(n)` out of the `r totalArticles` (`r reproAfter %>% filter(reproOutcome == F) %>% pull(percent)`) articles containing at least one major numerical discrepancy. The full breakdown of reproducibility outcomes after author contact are shown in Table \@ref(tab:reproTab). Aspects of `r as.english(insuffInfoErrors)` reproducibility checks could not be completed as there was insufficient information to proceed with the reanalysis. Availability of original analysis scripts appeared to provide some modest benefits in terms of reproducibility outcomes and time expenditure (see Supplementary Information \@ref(sup_results)). In no cases did the observed numerical discrepancies appear to be consequential for the conclusions stated in the original articles (see Supplementary Information \@ref(vignettes)).

@@

```{r reproTab}
data.frame("Target values fully reproducible?" = c("Yes", "Yes", "No", "No"),
           "Original author involvement?" = c("Yes", "No", "Yes", "No"),
           "n (%[95%CI])" = c(
             articleLevelSummary %>% 
               filter(finalOutcome == 'Reproducible\nwithout author\ninvolvement') %>%
               pull(fullReport),
             articleLevelSummary %>% 
               filter(finalOutcome == 'Reproducible\nwith author\ninvolvement') %>%
               pull(fullReport),
             articleLevelSummary %>% 
               filter(finalOutcome == 'Not reproducible\nwith author\ninvolvement') %>%
               pull(fullReport),
             articleLevelSummary %>% 
               filter(finalOutcome == 'Not reproducible\nwithout author\ninvolvement') %>%
               pull(fullReport))) %>% 
  kable(caption = "Final outcomes of reproducibility checks at the article level after original authors were contacted.", align = 'l', 
        col.names = c("Target values fully reproducible?",
                      "Original author involvement?",
                      "n (%[95%CI])"))
```

@@

Whilst the current findings are concerning, non-reproducibility is fortunately a solvable problem, at least in principle. As noted above, a major barrier to analytic reproducibility is low availability of research resources like data and code [@hardwicke2020a]. Some evidence suggests that journal policies which mandate data sharing are associated with increased data availability [@hardwicke2018a; @nuijten2017], apparently more so than voluntary badge schemes [@kidwell2016]. Sharing of analysis scripts may help by providing veridical documentation of the analysis process which is not well captured in verbal prose [@hardwicke2018a]. However, as with data sharing, mere availability will not be sufficient to confer the potential benefits of analysis scripts. The utility of scripts depends on a range of factors, including clarity, structure, and documentation, as well the programming language that is used and whether it is in a proprietary format or requires special expertise to understand. In the present study, script availability appeared to offer modest benefits in terms of reproducibility outcomes and time expenditure; however, as only six articles shared scripts[^psychsci_scripts], it is difficult to draw strong conclusions about their impact from this evidence alone.

@@

TEH and MCF designed the study. TEH, MB, KM, EH, MBN, BNP, BdM, BL, EJY, and MCF conducted reproducibility checks. TEH performed the data analysis. TEH and MCF wrote the manuscript. MBN provided feedback on the manuscript. All authors gave final approval for publication.

### 2020-06-17 manuscript/supplementary_information.Rmd
minor text edits after MN feedback

A precision analysis (not pre-registered and performed after study completion) shows the range of expected margin of error (confidence interval width) for the range of possible effect sizes given our sample size of 25 articles (Figure \@ref(fig:precisionCurve)). The maximum margin of error is `r maxMOE`.

@@

Outcome: Not fully reproducible (without author involvement)

>>>>

A precision analysis (not pre-registered and performed after study completion) shows the range of expected margin of error (confidence interval width) for the range of possible effect sizes (proportion) given our sample size of 25 articles (Figure \@ref(fig:precisionCurve)). The maximum margin of error is `r maxMOE`.

@@

Outcome: Not fully reproducible (with no author involvement)
